Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2017-12-29 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Correction

%%%%%%%%%%%%%
\subsection{3-D channel flows} \label{section:results_3D}

In this section, 3D simulations are introduced to investigate preferential flow pathways and finger growth through semi-pervious and impervious layers. In large scale geological formations preferential flow may allow fast contaminant transport and creates significant consequences for ground-water quality, and low recovery rate in hydrocarbon exploration.

\medskip
As mentioned in Section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} shows the initial permeability distribution for the simulations described in this section. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-slip conditions are applied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0) = 0.80$, which was displaced by a pure (wetting) fluid. The domain was designed to capture preferential flow pathways, as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

>>>>

%%%%%%%%%%%%%
\subsection{3-D channel flows} \label{section:results_3D} 
In geological formations, preferential flow may result in fast polutant transport (advection-diffusion-dispersion) after industrial spillage (therefore leading to contamination of groundwater) or low hydrocarbon recovery rates (in oil/gas exploration fields). In this Section, 3D simulations are introduced to investigate preferential flow pathways through semi-pervious and impervious geological layers.

As outlined in Section \ref{section:ViscousInstabilities}, viscous flow instabilities are characterised by spontaneous formation and growth of fluid elongations, called viscous fingers. Geometrical evolution in time and space of these elongations provides good indication of instabilities at the fluid front interface. Figure \ref{fig:3DChannel_PermMesh} shows the permeability distribution for the simulations conducted in this Section, where pure (wetting) fluid is added into the domain with uniform velocity of 2.5 cm.s$^{-1}$. No-flux boundary conditions are applied to all borders, except at the inlet (left-hand side) and outflow (right-hand side) faces. Initially, the domain was partially saturated with non-wetting fluid $\left(S_{nw}(t=0)=0.80\right)$, which was displaced by the wetting fluid as the simulation progress. The domain (with 368513 \PNDG[1]{1} tetrahedral element-pairs) was designed to naturally capture preferential flow pathways through channels of large absolute permeability (Fig.~\ref{fig:3DChannel_PermMesh}).% as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

No-flux boundary conditions are applied to all borders, except the right-hand (inlet) and the left-hand sides (outflow).. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-slip conditions are applied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0) = 0.80$, which was displaced by a pure (wetting) fluid. The domain was designed to capture preferential flow pathways, as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.
